---
title: "Data Cleaning and Preprocessing"
author: "Zach Weber"
date: "March 27, 2019"
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
To prepare our training data for model creation through pruning and feature engineering.

# Loading in Libraries and Data

First, lets load in some libraries that may be useful for our purposes
```{r, results='hide', message=FALSE, warning=FALSE}
library(dplyr, quietly=TRUE)
library(ggplot2, quietly=TRUE)
```

Next we can load in our training data 
```{r}
# generate path to training data
data.dir <- "datasets/"
train.fpath <- paste0(data.dir,"train.csv")

# load in training data
train.dat <- read.table(train.fpath, sep=",", header=TRUE)
```

# Exploring Features
Lets take a look at what features we have. We know that our goal is to predict the price of the home which is encoded in the "SalePrice" variable, but what predictors do we have acceess to and what type of variable are they?

```{r}
# get the number of predictors 
length(colnames(train.dat))-1 # columns minus SalePrice

# get the colnames of the predictors
featnames <- colnames(train.dat)[which(colnames(train.dat)!="SalePrice")]
featnames
```

It might be helpful to see which predictors are continuous and which predictors are categorical in this list. We can get a rough split by querying the data type of the column
```{r}
# split features into numeric and categorical (roughly)
numericFeats <- featnames[unlist(lapply(train.dat[,featnames], is.numeric))] 
categoricalFeats <- featnames[!featnames %in% numericFeats]

# display numeric features 
numericFeats

# display categorical features
categoricalFeats
```


## Pruning Predictors
We also should be interested in which features have missing data in both the numeric and categorical
type. Because missing data imputation is not totally within the scope of this class we would probably
remove all sets of predictors with missing data. We can start by generating filters for columns 
with missing data.
```{r}
# generate filters for missing data
num.NA.filt <- unlist(lapply(train.dat[,numericFeats],
                             function(x) any(is.na(x))))
cat.NA.filt <- unlist(lapply(train.dat[,categoricalFeats],
                             function(x) any(is.na(x))))
```

Lets summarize the findings when we apply the filter (print statements not shown)
```{r, echo=FALSE}
# number numeric features without NAs + list of vars with NA
cat(sprintf("Number of Numeric Features: %i\n", length(numericFeats)))
cat(sprintf("Number of Numeric Features w/o NAs: %i\n", 
             length(numericFeats[!num.NA.filt])))
cat("Features with NAs: ", numericFeats[num.NA.filt], "\n")

# number of categorical features without NAs + list of vars with NA 
cat(sprintf("Number of Categorical Features: %i\n", length(categoricalFeats)))
cat(sprintf("number of Categorical Features w/o NAs: %i\n",
            length(categoricalFeats[!cat.NA.filt])))
cat("Features with NAs:\n")
categoricalFeats[cat.NA.filt]



```

Now that we've identified those features with missing data, we can remove them from the data frame
to produce a new pruned dataset. We should have 34 + 27 = 61 features remaining on which to predict. The total number of columns should be 61 + 1 =62 for the features and the _SalePrice_ variable

```{r}
# get the column names to be removed
columns.to.keep <- c(categoricalFeats[!cat.NA.filt],
                     numericFeats[!num.NA.filt],
                     "SalePrice")
pruned.data <- train.dat[,columns.to.keep]
dim(pruned.data)
```

Everything here looks good. We wont directly create dummy variables here since R handles factors pretty intuitively. 

## Saving Our Pruned Data

We're going to save our pruned data table in two different formats. We'll save it first as a __.csv__ file and then we'll also save it as a __.rda__ file for easier reading.

```{r}
file.prefix <- "pruned_data"

# save as .csv
write.table(x=pruned.data,
            file=paste(data.dir,file.prefix,".csv"),
            sep=",",
            row.names=FALSE,
            quote=FALSE)

# save as .rda
save(x=pruned.data, file=paste0(data.dir, file.prefix, ".rda"))
```



