---
title: "EDA"
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(DataExplorer)
```

Exploratory data analysis. There are 62 variables. Approximately half of them are categorical, and the other half is of integer values. The histograms show that the variables follow different distributions, often skewed and/or bimodal.
```{r}
train <- read.table("/Users/polina/Projects/house-prices-regression/datasets/pruned_data.csv", sep=",", header=TRUE)
plot_str(train)
plot_histogram(train)
```

Multivariate analysis. The correlation plot shows that variables such as OverallQual, GrLivArea, FullBath, TotRmsAbvGrd, GarageArea, GarageCars, and SalePrice are more strongly positively correlated with other variables, while EnclosedPorch is the only one that is overall negatively correlated.
```{r}
plot_correlation(train, type = 'continuous','Review.Date')
```


